---
title: "Merge Collaborative Projects"
output: html_notebook
root.dir: 
---
Merge collaborative projects based on title of grant. Add all funding and place on one grant. This is because the Abstracts for all proposals are the same distorting topic analysis. Some project titles on collaborative grants are however slightly different so some matching may need to be by hand.

# set the root directory for analysis
# setwd("~/Documents/TextMiningR/NSF-AST/")
#setwd("~/Documents/papers/NSFThemeAnalysis2018/RTrendAnalysis/")

#tkmessageBox(message = "Select a tab delimited text file.")
#filename <- tclvalue(tkgetOpenFile(initialdir=".")) # 


```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/heidorn/Desktop/NSF201020152020')
```
```{r}
#setwd("/Users/heidorn/Desktop/NSF201020152020")

filename = file.choose()
NSF_Funding <- read.csv(filename, header=T, encoding = "UTF-8", 
                        stringsAsFactors=FALSE) # comma delimited
# remove empty columns
empty_columns <- colSums(is.na(NSF_Funding) | NSF_Funding == "") == nrow(NSF_Funding)
# Change data type to number for award amount
NSF_Funding$AwardedAmountToDate <- as.numeric(gsub('[$,]', '', NSF_Funding$AwardedAmountToDate))
#NSFTemp <- NSF_Funding[, !empty_columns]
#Make a subdirectory if needed to save the results
topicDir <- tools::file_path_sans_ext(filename)
ifelse(!dir.exists(topicDir), dir.create(topicDir), "")

# drop records with empty abstracts. These are usually intergovernmental agreements
NSF_Funding <- NSF_Funding[!NSF_Funding$Abstract=="", ]
```

```{r}
# Create a triangular matrix of similarity.
# Nested loop start with 1 - N
library(dplyr)
#https://cran.r-project.org/web/packages/stringdist/stringdist.pdf
install.packages("stringdist")
library(stringdist)
Mx <- stringdistmatrix(tolower(NSF_Funding$Title), tolower(NSF_Funding$Title), method= c("lcs"))
# Too much data to be helpful in this visualization 
#plot(hclust(as.dist(Mx)))
#serialization inspired by: https://stackoverflow.com/questions/27892100/distance-matrix-to-pairwise-distance-list-in-r
# Use the AwardNumber as the labels
dimnames(Mx) <- dimnames(Mx) <- list(NSF_Funding$AwardNumber, NSF_Funding$AwardNumber) 
#  Paste together the AwardNumber to matrix
xy <- t(combn(colnames(Mx), 2))
MxSim <- data.frame(xy, dist=Mx[xy])
# Now sort by similarity and AwardNumber.
MxSim <- MxSim[ with(MxSim, order(MxSim$dist, MxSim$X1)), ]
# Shorten the list to the most similar
ShortMxCol <- MxSim[ which(MxSim$dist < 25), ]
# add the Project Titles 
# Add the title for the left AwardNumber
ShortMxCol <- merge(x = ShortMxCol, y = select(NSF_Funding, c(AwardNumber, Title)), by.y = c("AwardNumber"), by.x = "X1")
# Change the name of the "Title" Column to "Title 1"
colnames(ShortMxCol)[which(names(ShortMxCol) == "Title")] <- "Title 1"
# Add the title for the right AwardNumber
ShortMxCol <- merge(x = ShortMxCol, y = select(NSF_Funding, c(AwardNumber, Title)), by.y = c("AwardNumber"), by.x = "X2")
write.csv(ShortMxCol, paste(topicDir,"/SimilarTitles.csv", sep=""))
```


```{r}
# Repeat calculations for abstracts
#_____________________________________________________
# remove html tags in the form <anything>
# remove where the word is a number
# if there are any "words" with embedded periods Eg "John.P.Doud, replace the period with a space.
# remove "'s"
# fix eg trichoptera:hydropsychidae 
NSF_Funding$Abstract <- gsub("<[^>]+>", " ", 
                        gsub("\\d", "", 
                        gsub("\\.", " ", 
                        gsub("'s", "", 
                        gsub(":", " ", NSF_Funding$Abstract)))))

# if you get errors from multi-byte characters you may need to run: iconv  -f UTF-8 -t ISO-8859-1 -c Awardsfile.csv > Awardsfile.csv
# make a shortened abstract of 100 characters
NSF_Funding$ShortAbstract <- tolower(substr(NSF_Funding$Abstract, 1, 100))

MxAbstract <- stringdistmatrix(NSF_Funding$ShortAbstract,     
                               NSF_Funding$ShortAbstract, method= c("cosine"))
# Too much data to be helpful in this visualization 
#plot(hclust(as.dist(Mx)))
#serialization inspited by: https://stackoverflow.com/questions/27892100/distance-matrix-to-pairwise-distance-list-in-r
# Use the AwardNumber as the labels
dimnames(MxAbstract) <- dimnames(MxAbstract) <- list(NSF_Funding$AwardNumber, NSF_Funding$AwardNumber) 
#  Paste together the AwardNumber to the matrix
xyAbstract <- t(combn(colnames(MxAbstract), 2))

MxSimAbstract <- data.frame(xyAbstract, dist=MxAbstract[xyAbstract])
# Now sort by similarity and AwardNumber.
MxSimAbstract <- MxSimAbstract[ with(MxSimAbstract, order(MxSimAbstract$dist, MxSimAbstract$X1)), ]
# Shorten the list to the most similar
ShortMxColAbstract <- MxSimAbstract[ which(MxSimAbstract$dist < .01), ]
# add the Project Abstract 
# Add the title for the left AwardNumber
ShortMxColAbstract <- merge(x = ShortMxColAbstract, y = 
                              select(NSF_Funding, c(AwardNumber, ShortAbstract, 
                                                    ProgramElementCode.s.)), 
                              by.y = c("AwardNumber"), by.x = "X1")
# Change the name of the "Abstract" Column to "Abstract 1"
colnames(ShortMxColAbstract)[which(names(ShortMxColAbstract) == "ShortAbstract")] <- "SAbstract 1"
colnames(ShortMxColAbstract)[which(names(ShortMxColAbstract) == "ProgramElementCode.s.")]  <- "ProgramElementCode1"
# Add the title for the right AwardNumber
ShortMxColAbstract <- merge(x = ShortMxColAbstract, y = 
                              select(NSF_Funding, c(AwardNumber, ShortAbstract, 
                                                    ProgramElementCode.s.)), 
                            by.y = c("AwardNumber"), by.x = "X2")
colnames(ShortMxColAbstract)[which(names(ShortMxColAbstract) == 
                                     "ProgramElementCode.s.")]  <- "ProgramElementCode2"
write.csv(ShortMxColAbstract, paste(topicDir,"/SimilarAbstracts.csv", sep=""))
```


# Marge all records remaining in the Similarity file.
# Any records that should not be there should have been removed by a manual process.
ShortMxColWOCollab <- read.csv(paste(topicDir,"/SimilarTitles.csv", sep=""), header=TRUE, encoding = "UTF-8", stringsAsFactors=FALSE) # comma delimited
# Read through main file line by line... see if matches exists in the Title column with similarity < 25. If so merge the $amount and the ProgramElementCode(s) and ProgramReferenceCode(s) removing duplicates.'
```{r}

```

```{r}
Collaborationsdf = subset(ShortMxColAbstract, 
                          select = c(X1, X2, ProgramElementCode1, ProgramElementCode2) )
# Merge Program element codes, group by X1
Collaborationsdf <- merge(x = Collaborationsdf, y = 
                              select(NSF_Funding, c(AwardNumber,AwardedAmountToDate)), 
                            by.y = c("AwardNumber"), by.x = "X1")
#rename AwardedAmountToDate to AwardedAmountToDateX1
colnames(Collaborationsdf)[which(names(Collaborationsdf) 
                                 == "AwardedAmountToDate")] <- "AwardedAmountToDateX1"
# Merge Program element codes, group by X2
Collaborationsdf <- merge(x = Collaborationsdf, y = 
                              select(NSF_Funding, c(AwardNumber,AwardedAmountToDate)), 
                            by.y = c("AwardNumber"), by.x = "X2")
#rename AwardedAmountToDate to AwardedAmountToDateX2
colnames(Collaborationsdf)[which(names(Collaborationsdf) 
                                 == "AwardedAmountToDate")] <- "AwardedAmountToDateX2"
Collaborationsdf <- Collaborationsdf[, c(2, 1, 3, 4, 5, 6)] # leave the row index blank to keep all rows
library(dplyr)
# csum will have sum of the collaborative grants plus the starting grant
Collaborationsdf <- Collaborationsdf %>% 
  group_by(X1) %>% 
  mutate(csum = sum(AwardedAmountToDateX2) + AwardedAmountToDateX1) #%>%
 # AwardedAmountToDateX2 = 0

# for debugging only Collaborationsdf <- subset(Collaborationsdf, select = -c(csum) )
# Make copy of NSF_Funding with all funding records
NSFRecordsSkipDF <- NSF_Funding
#Add a true/false column to indicate that a field should be skipped on output (because of merged records)
NSFRecordsSkipDF['Skip'] <- FALSE
#scan through each records in the dataframe that is not marked SKIP
collaborators <- function(ID) {
  #select X2 for all rows with ID as the X1 project ID
  leftrepeats <- Collaborationsdf[Collaborationsdf$X1 == '1010009', "X2"]
}
#check the ShortMxColAbstract AwardNumber1  if it matches, just keep it get the awardnumber2 from the #similarity matrix. search the NSFFunding for the Amount and add it to the Amount in the # NSFRecordSkipDF, concatinate the element codes and remove duplicates. make the record as SKIP.
#get all AwardNumber1 where the AwardNumber matches ShortMxColAbstract$AwardNumber2
# for each of these add the amount and concat the element codes and make as skip.# eliminate all records marked as skip

# do a check by comparing the sum of amounts in the original data and the new colaboration combined data
# 
```

